Assigning Scouts to Optimal Patrols

### 2016-06-08 draft1.tex
General edits
Added some details

\begin{abstract}
Scoutmaster Gene Bruce of the Boy Scouts of America for Troop 407 in Kent, Washington wants to find an arrangement of 13 new scouts into two patrols. The cohesiveness and overall quality experience of a patrol is affected by adverse relationships and existing friendships. By using anonymous data from scouts and parents, a brute force strategy was employed to find an optimal arrangement of scouts that avoids severe conflict while maximizing pre-existing preferences. 
\end{abstract}

@@

We aim to answer the particular instance of this problem provided by our community partner, that of arranging 13 scouts into two troops that avoids severe conflicts while promoting positive relationships between scouts. Furthermore, we want to determine if we can build a model to determine optimal patrol arrangements for any number of scouts so that the Scoutmaster can use this model for future years. For 13 scouts, a grouping into one patrol is too large, while three patrols are too small, and need not be considered. But in general it is possible to calculate all possible partitions with size constraints on the partitions for an arbitrary number of scouts\footnote{Desquelles, P. "Calculation of the Number of Partitions with Constraints on the Fragment Size." Phys. Rev. C Physical Review C 65.3 (2002)}. Additionally, this model could be applied to other situations in which arranging a number of individuals into groups where interpersonal dynamics are necessary to consider - such as teams within a company or military troops.
The 13 different recruited scouts are Brandon, Cameron, Christian, Colby, Daniel, Darwin, Evan, Jake, Jordan, Nathan, Patrick, Timmy and Tommy. After the Scoutmaster collected the likes and dislikes of each scout, he presented the data as follows: 
\begin{figure}[h]
	\centering
	\includegraphics[scale=0.60]{graph}
\end{figure}

@@

\section*{Simplifications}
Ultimately, our goal is to find groups for the scouts that will allow them to enjoy their time scouting to the fullest. Modeling interpersonal relations by likes and dislikes is, of couse, a simplification. But the difficulty in modeling the intricacies of human relationships demands it. In addition, our process will have to be repeated each year for every new batch of scouts, and would perhaps be too burdensome to warrant the effort if our process attempted to delve too far into the psyche.

Another simplification is to assume that the quality of a group can be determined by the quality of relationships between unique pairs either inside or outside the group - in other words, the whole is equal to the sum of the parts. For example, consider the case where scouts A, B and C individually like each other when they are alone with one of their group, but don't like being in a group together. This type of epiphenomenon won't be considered.

\section*{Mathematical Model}
To model the relationships between scouts, we use a mathematical graph. We let the scouts be vertices in our graph. A directed edge between Scout A and Scout B has a weight proportional to how much Scout A likes Scout B. We would then like to find a partition of the vertices in the graph that either minimizes or maximizes an objective function given constraints on the sizes of subsets generated by our partition. This results in a complete, directed graph on 13 vertices having edge weights 0, 1, or -1, corresponding to indifference, approval, or disapproval, respectively. Scoutmaster Bruce specifically would like two patrols from the 13 scouts - one of size 6, the other of size 7. The total number of partitions given these constrants is $\binom{13}{6} = \binom{13}{7} = 1716$. This number is reasonable enough that we may calculate the total loss of every valid grouping and select the partition with the lowest or highest score depending on the objective functionn.

@@

Of the objective functions we chose to use, it seems like MinCut and HybridCut are the best performing. FriendCut performs almost as well as HybridCut, except for one worse in the EnemyCut objective. EnemyCut does poorly since finding a partition with 0 intra-patrol negative weight edges is easy in this example, and so EnemyCut settles for the first suboptimal (in terms of the other objective functions) solution it finds. AwkwardCut (+1) arrived at the same partition as MinCut (not altogether surprising, but informative, since even with the additional +1 penalty for placing Brandon and Tommy in the same patrol it decided it would be best off to do so anyways). But changing the ``awkward'' penalty to +2, the same penalty as mutual dislike in EnemyCut, we get a completely different partition.

@@

	\item As mentioned in the previous point, Tommy dislikes Brandon. Trading Tommy in $P1$ for Jordan in $P2$ seems promising, But we have overlooked the fact that Colby and Jordan are our NA pair, and cannot be in the same patrol under any circumstances. Trading Colby for Jordan is another option, but Colby will be missed by Darwin, Evan, Cameron and Tommy (In other words, the whole of P1 minus Brandon).
\end{enumerate}

@@

\section*{Improvements}
While our algorithm demonstrably works well on smaller groups of 13 people, we do not expect our brute force solution to be tractable on larger groups wherein greater than 2 partitions are required. As a possible scenario, consider a company of 150 people to be divided into fixed-size teams of 10 individuals each. There are $150! / (10!)^{15}$ possible partitions, a number proportional to $10^{164}$ - meaning we could instead use that computational time to enumerate the number of atoms in the universe\ldots a tredecillion $(10^{78})$ times. Knowing this, perhaps it won't surprise the reader that this particular graph partitioning problem in general graphs is NP-complete - though there do exist approximation algorithms\footnote{B. W. Kernighan, S. Lin, \textit{An Efficient Heuristic Procedure for Partitioning Graphs}. Bell System Technical Journal. \textbf{49} (1970), 291--307. doi: 10.1002/j.1538-7305.1970.tb01770.x}.
As scout enrollment increases, it becomes necessary to consider scenarios where we have both a large number of scouts in need of assignment and multiple patrols to choose from. Our algorithm could be extended to use brute force to find a globally optimal solution when it is deemed computationally feasible and to use an approximating algorithm to find a locally optimal solution otherwise.

\section*{Conclusions}
Assigning individuals to compatible groups is hard (computationally speaking). When the group is small enough - as it is here - globally optimal solutions may be arrived at by checking all feasible solutions. But the amount of computation time needed to arrive at an answer grows exponentially with the number of individuals and groups. For example, finding an optimal partition of 13 scouts into \textit{three} groups would take over 50 times as long as finding the optimal partition of 13 scouts into two groups (90,090 feasible combinations to check versus 1,716 in the two patrol case).

@@

\bibitem {B} Desquelles, P. "Calculation of the Number of Partitions with Constraints on the Fragment Size." Phys. Rev. C Physical Review C 65.3 (2002)

@@

>>>>

\begin{abstract}
Scoutmaster Gene Bruce of the Boy Scouts of America for Troop 407 in Kent, Washington wants to find an arrangement of 13 new scouts into two patrols. The cohesiveness and overall quality experience of a patrol is affected by conflicting relationships and existing friendships. If enough adverse encounters occur in one patrol, the scouts may decide to drop out from the organization all together, missing out on a potentially great experience and community of friends. Thus our project focuses on helping the Scoutmaster find an optimal arrangement of the new scouts into tw patrols By using anonymous data from scouts and parents, a brute force strategy was employed to find an optimal arrangement of scouts that avoids severe conflict while maximizing pre-existing preferences. 
\end{abstract}

@@

We aim to answer the particular instance of this problem provided by our community partner, that of arranging 13 scouts into two troops that avoids severe conflicts while promoting positive relationships between scouts. Furthermore, we want to determine if we can build a model to determine optimal patrol arrangements for any number of scouts so that the Scoutmaster can use this model for future years. For 13 scouts, a grouping into one patrol is too large, while three patrols are too small, and need not be considered. But in general it is possible to calculate all possible partitions with size constraints on the partitions for an arbitrary number of scouts\footnote{Desquelles, P. "Calculation of the Number of Partitions with Constraints on the Fragment Size." Phys. Rev. C Physical Review C 65.3 (2002)}. Additionally, this model can be applied to elementary school students to promote supportive and enthusiastic groups for better learning, ensuring cooperation and harmony in military troops for mission success and survival (\textit{Personal Interviews}), and other situations where interpersonal dynamics are necessary.

The 13 different recruited scouts are Brandon, Cameron, Christian, Colby, Daniel, Darwin, Evan, Jake, Jordan, Nathan, Patrick, Timmy and Tommy. After the Scoutmaster collected the likes and dislikes of each scout, he presented the data as follows: 
\begin{figure}[h]
	\centering
	\includegraphics[scale=0.60]{graph}
\end{figure}

@@

\section*{Simplifications}
Ultimately, our goal is to find groups for the scouts that will allow them to enjoy their time scouting to the fullest. Modeling interpersonal relations by likes and dislikes is, of course, a simplification. But the difficulty in modeling the intricacies of human relationships demands it. In addition, our process will have to be repeated each year for every new batch of scouts, and would perhaps be too burdensome to warrant the effort if our process attempted to delve too far into the psyche.

Another simplification is to assume that the quality of a group can be determined by the quality of relationships between unique pairs either inside or outside the group - in other words, the whole is equal to the sum of the parts. For example, consider the case where scouts A, B and C individually like each other when they are alone with one of their group, but do not like being in a group together. This type of epiphenomenon will not be considered.

\section*{Mathematical Model}
To model the relationships between scouts, we use a mathematical graph. We let the scouts be vertices in our graph. A directed edge between Scout A and Scout B has a weight proportional to how much Scout A likes Scout B. We would then like to find a partition of the vertices in the graph that either minimizes or maximizes an objective function given constraints on the sizes of subsets generated by our partition. This results in a complete, directed graph on 13 vertices having edge weights 0, 1, or -1, corresponding to indifference, approval, or disapproval, respectively. Scoutmaster Bruce specifically would like two patrols from the 13 scouts - one of size 6, the other of size 7. The total number of partitions given these constrants is $\binom{13}{6} = \binom{13}{7} = 1716$. This number is reasonable enough that we may calculate the total loss of every valid grouping and select the partition with the lowest or highest score depending on the objective function.

@@

Of the objective functions we chose to use, it seems as though MinCut and HybridCut are the best performing. FriendCut performs almost as well as HybridCut, except for one worse in the EnemyCut objective. EnemyCut does poorly since finding a partition with 0 intra-patrol negative weight edges is easy in this example, and so EnemyCut settles for the first suboptimal (in terms of the other objective functions) solution it finds. AwkwardCut (+1) arrived at the same partition as MinCut (not altogether surprising, but informative, since even with the additional +1 penalty for placing Brandon and Tommy in the same patrol it decided it would be best off to do so anyways). But changing the ``awkward'' penalty to +2, the same penalty as mutual dislike in EnemyCut, we get a completely different partition.

@@

	\item As mentioned in the previous point, Tommy dislikes Brandon. Trading Tommy in $P1$ for Jordan in $P2$ seems promising, but we have overlooked the fact that Colby and Jordan are our NA pair, and cannot be in the same patrol under any circumstances. Trading Colby for Jordan is another option, but Colby will be missed by Darwin, Evan, Cameron and Tommy (In other words, the whole of P1 minus Brandon).
\end{enumerate}

@@

\section*{Improvements}
While our algorithm demonstrably works well on smaller groups of 13 people, we do not expect our brute force solution to be tractable on larger groups wherein greater than 2 partitions are required. As a possible scenario, consider a company of 150 people to be divided into fixed-size teams of 10 individuals each. There are $150! / (10!)^{15}$ possible partitions, a number proportional to $10^{164}$ - meaning we could instead use that computational time to enumerate the number of atoms in the universe\ldots a tredecillion $(10^{78})$ times. Knowing this, perhaps it will not surprise the reader that this particular graph partitioning problem in general graphs is NP-complete - though there do exist approximation algorithms\footnote{B. W. Kernighan, S. Lin, \textit{An Efficient Heuristic Procedure for Partitioning Graphs}. Bell System Technical Journal. \textbf{49} (1970), 291--307. doi: 10.1002/j.1538-7305.1970.tb01770.x}.
As scout enrollment increases, it becomes necessary to consider scenarios where we have both a large number of scouts in need of assignment and multiple patrols to choose from. Our algorithm could be extended to use brute force to find a globally optimal solution when it is deemed computationally feasible and to use an approximating algorithm to find a locally optimal solution otherwise.

\section*{Conclusions}
Assigning individuals to compatible groups is hard (computationally speaking). When the group is small enough - as it is here - globally optimal solutions may be arrived at by checking all feasible solutions. However, the amount of computation time needed to arrive at an answer grows exponentially with the number of individuals and groups. For example, finding an optimal partition of 13 scouts into \textit{three} groups would take over 50 times as long as finding the optimal partition of 13 scouts into two groups (90,090 feasible combinations to check versus 1,716 in the two patrol case).

@@

\bibitem {B} Desquelles, P. ``Calculation of the Number of Partitions with Constraints on the Fragment Size." Phys. Rev. C Physical Review C 65.3 (2002)

@@

\bibitem {D} Mendoza, Veronica, ``Mabel Paine Elementary School classroom helper." \textit{Personal Communication}.

\bibitem {E} Lin, Kevin. ``Air Force ROTC Cadet in the 910th Cadet Wing." \textit{Personal Communication}.
